Lifting CDCL to Numerical Domains

### 2016-09-27 domains.tex
more on program model

\section{Abstract Domains}
\textbf{Programs:}
%
A {\em program} $\mathcal{P}$ is formally defined as follows.
\[
\begin{array}[t]{@{}lll}
\mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
\end{array}
\]
Consider sets of expressions $Exp$ and Boolean expressions $BExp$
over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
can take numeric values in $Val$.  A procedure is denoted as 
$\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
is an assignment, conditional, sequential concatenation or a loop.
\rmcmt{post}
\subsection{Interval Abstract Domain}
The {\em Interval} abstract domain is a lattice 
$ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
$ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
the set of intervals of type $[l,u]$ over numeric data 
type with $l \leq u$. The least element is $\bot$ and the 
greatest element is $\top$ which maps all variables to their
minimum ($min$) and maximum ($max$) values.  An interval 
$\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
partial order $\sle_{I}$ over elements in the set $Itv$ is 
given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
The galois connection between the concrete environment and intervals is 
given as follows.
\[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
\[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
For a program with $N$ variables, there are a total of 
$2*N$ intervals. The intervals restrict the set of values 
taken by program variables at each program point, starting 
from $[-\infty, \infty]$.

@@

The abstract transformer, $apost$, captures the effect of different program 
statements in the abstract domain. The transformer is precise for octagonal 
assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
as shown below.
\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \]  
\[apost_{x:=y+z}(b) = \langle \top \rangle \]

>>>>

\section{Program and Abstract Domains}
\subsection{Concrete Semantics}
\textbf{Programs:}
%
A {\em program} $\mathcal{P}$ is formally defined as follows.
\[
\begin{array}[t]{@{}lll}
\mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
\end{array}
\]
Consider sets of expressions $Exp$ and Boolean expressions $BExp$
over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
can take numeric values in $Val$.  A procedure is denoted as 
$\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
is an assignment, conditional, sequential concatenation or a loop. \\

\textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
where $Loc$ is the set of locations with an unique start 
location $(init)$ and error location $(err)$, $E$ is the set 
of control-flow edges that are labelled with the set 
$lbl \in Statement$.  For the purpose of illustration, we 
assume that all the procedures are inlined.

\textbf{Concrete Domain:} The {\em concrete} domain is a lattice 
of concrete environments, $Env = Var \rightarrow Val$, and is 
defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
A transformer, $post_{stmt}$, of concrete domain defines 
the effect of a statement $stmt$ on the concrete domain, 
$post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.

A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
$0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
if $\sigma_{i+1} \in post_{i}(\sigma_{i})$. \\

\subsection{Interval Abstract Domain}
The {\em Interval} abstract domain is a lattice 
$ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
$ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
the set of intervals of type $[l,u]$ over numeric data 
type with $l \leq u$. The least element is $\bot$ and the 
greatest element is $\top$ which maps all variables to their
minimum ($min$) and maximum ($max$) values.  An interval 
$\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
partial order $\sle_{I}$ over elements in the set $Itv$ is 
given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
$\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
The galois connection between the concrete environment and intervals is 
given as follows.
\[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
\[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
For a program with $N$ variables, there are a total of 
$2*N$ intervals. The intervals restrict the set of values 
taken by program variables at each program point, starting 
from $[-\infty, \infty]$.

@@

The abstract transformer, $apost_{stmt}$, captures the effect of different program 
statements in the abstract domain. The transformer is precise for octagonal 
assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
as shown below.
\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \]  
\[apost_{x:=y+z}(b) = \langle \top \rangle \]
